---
title: "Getting Started with AgentLib"
author: "Alexander Alexeyuk"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with AgentLib}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Set to FALSE to avoid API calls during package build
)
```

# Introduction

AgentLib is a comprehensive R package for building advanced AI agents using the ellmer framework. This vignette will guide you through the key features and demonstrate how to create and use different types of agents.

## Installation

```{r setup}
# Install from GitHub (development version)
# devtools::install_github("A1exanderAlexeyuk/AgentLib")

library(AgentLib)
```

## Basic Agent Usage

### Creating Your First Agent

The `BaseAgent` class provides the foundation for all AI agents in AgentLib:

```{r basic-agent}
# Create a basic agent
agent <- BaseAgent$new(
  provider = "openai/gpt-4",
  agent_name = "MyAssistant",
  system_prompt = "You are a helpful AI assistant that provides clear, concise answers.",
  logging_enabled = TRUE
)
```

### Having Conversations

Once created, you can interact with your agent using the `ask()` method:

```{r conversation}
# Simple conversation
response <- agent$ask("What is machine learning?")
print(response)

# Ask follow-up questions
response2 <- agent$ask("Can you give me a practical example?")
print(response2)
```

### Using Built-in Tools

All agents come with useful built-in tools:

```{r tools}
# Get current time
time_response <- agent$ask("What time is it in Tokyo?")
print(time_response)

# Perform calculations
calc_response <- agent$ask("Calculate the square root of 256")
print(calc_response)

# Use memory functions
agent$ask("Remember that my favorite color is blue")
memory_response <- agent$ask("What is my favorite color?")
print(memory_response)
```

## Specialized Agents

### Medical Writer Agent

The `MedicalWriterAgent` is specialized for medical writing tasks:

```{r medical-agent}
# Create medical writing agent
medical_agent <- MedicalWriterAgent$new(
  provider = "openai/gpt-4",
  agent_name = "MedicalExpert"
)

# Create clinical protocol
protocol <- medical_agent$create_protocol(
  study_title = "Efficacy of Novel Antihypertensive Agent",
  indication = "Essential Hypertension",
  study_design = "Randomized, Double-blind, Placebo-controlled Trial",
  primary_endpoint = "Change in systolic blood pressure from baseline",
  sample_size = "300 patients"
)
print(protocol)
```

#### Medical Writing Features

```{r medical-features}
# Generate regulatory summary
reg_summary <- medical_agent$create_regulatory_summary(
  drug_name = "Cardioprotect-XYZ",
  indication = "Hypertension",
  submission_type = "NDA",
  key_findings = "Significant BP reduction with good safety profile"
)

# Create medical abstract
abstract <- medical_agent$create_abstract(
  title = "Novel ACE Inhibitor in Hypertensive Patients",
  study_type = "Randomized Controlled Trial",
  results_summary = "30% reduction in systolic BP, 25% reduction in diastolic BP",
  conclusion = "Demonstrates superior efficacy with good tolerability",
  word_limit = 250
)
```

### Data Analytics Agent

The `DataAnalyticsAgent` specializes in data analysis and machine learning:

```{r analytics-agent}
# Create analytics agent
analytics_agent <- DataAnalyticsAgent$new(
  provider = "openai/gpt-4",
  agent_name = "DataScientist"
)

# Generate and load sample data
data_summary <- analytics_agent$load_dataset(
  dataset_name = "sales_data",
  description = "Monthly sales performance data for analysis"
)
print(data_summary)
```

#### Analytics Workflow

```{r analytics-workflow}
# Perform statistical analysis
analysis <- analytics_agent$analyze_data(
  dataset_name = "sales_data",
  analysis_type = "descriptive",
  variables = "revenue, units_sold",
  hypothesis = "Revenue correlates with units sold"
)

# Create visualizations
viz <- analytics_agent$create_visualization(
  dataset_name = "sales_data",
  plot_type = "scatter",
  variables = "units_sold, revenue",
  title = "Sales Performance Analysis"
)

# Build predictive model
model <- analytics_agent$build_model(
  dataset_name = "sales_data",
  model_type = "linear_regression",
  target_variable = "revenue",
  feature_variables = "units_sold, region"
)

# Generate comprehensive report
report <- analytics_agent$generate_report(
  analysis_name = "Sales Performance Analysis",
  datasets = "sales_data", 
  key_findings = "Strong correlation between units sold and revenue",
  recommendations = "Focus marketing efforts on high-performing regions"
)
```

## Agent Factory

For quick agent creation, use the agent factory:

```{r factory}
# Create different agents quickly
base_agent <- create_agent("base", provider = "openai/gpt-4")
medical_agent <- create_agent("medical", provider = "anthropic/claude-3")
analytics_agent <- create_agent("analytics", provider = "openai/gpt-4")

# All agents are immediately ready to use
response1 <- base_agent$ask("Hello!")
response2 <- medical_agent$ask("What is pharmacovigilance?")
response3 <- analytics_agent$ask("Explain cross-validation")
```

## Custom Tools

Extend agent capabilities with custom tools:

```{r custom-tools}
# Create a sentiment analysis tool
sentiment_tool <- ellmer::tool(
  function(text) {
    # Simple sentiment analysis implementation
    positive_words <- c("good", "great", "excellent", "amazing", "wonderful")
    negative_words <- c("bad", "terrible", "awful", "horrible", "disappointing")
    
    text_lower <- tolower(text)
    pos_count <- sum(sapply(positive_words, function(w) grepl(w, text_lower)))
    neg_count <- sum(sapply(negative_words, function(w) grepl(w, text_lower)))
    
    if (pos_count > neg_count) return("Positive")
    if (neg_count > pos_count) return("Negative")
    return("Neutral")
  },
  name = "analyze_sentiment",
  description = "Analyze the sentiment of text input",
  arguments = list(
    text = ellmer::type_string("Text to analyze for sentiment")
  )
)

# Register with agent
agent$register_tool(sentiment_tool)

# Use the custom tool
result <- agent$ask("Analyze sentiment: 'This product is amazing and I love it!'")
print(result)
```

## Advanced Features

### Batch Processing

Process multiple prompts efficiently:

```{r batch}
# Prepare multiple prompts
prompts <- c(
  "What is artificial intelligence?",
  "Explain machine learning briefly",
  "What are neural networks?",
  "How does deep learning work?"
)

# Process in batch
results <- batch_process(
  agent = agent,
  prompts = prompts,
  delay_seconds = 1,
  save_results = TRUE
)

# Results include responses, timestamps, and cost information
str(results[[1]])
```

### Agent Comparison

Compare how different agents handle the same question:

```{r comparison}
# Create comparison
comparison <- compare_agents(
  agents = list(base_agent, medical_agent),
  prompt = "What are the key components of a clinical trial?",
  agent_names = c("BaseAgent", "MedicalAgent")
)

# View results
print(comparison$summary)
```

### Performance Monitoring

Monitor agent performance and resource usage:

```{r monitoring}
# Validate agent configuration
validation <- validate_agent(agent, run_tests = TRUE)
print(validation$summary)

# Get agent status
status <- agent$get_status()
print(status)

# Export configuration for reproducibility
config <- export_agent_config(agent, include_tools = TRUE)
str(config, max.level = 2)
```

### Conversation Management

Manage conversation history and context:

```{r conversation-mgmt}
# Save conversation to file
agent$save_conversation("my_session.rds")

# Get conversation history
history <- agent$get_history()
length(history)

# Get token usage and costs
tokens <- agent$get_tokens()
cost <- agent$get_cost()

# Clear history when needed
agent$clear_history()
```

## Error Handling and Best Practices

### Robust Error Handling

```{r error-handling}
# Wrap agent calls in error handling
tryCatch({
  response <- agent$ask("Complex query that might fail")
  print(response)
}, error = function(e) {
  cat("Error occurred:", e$message, "\n")
  # Handle error appropriately
})
```

### Best Practices

1. **Set appropriate context lengths** based on your use case
2. **Use descriptive agent names** for better logging and debugging
3. **Enable logging** for production applications
4. **Monitor costs** especially with large-scale deployments
5. **Validate agents** before putting them into production
6. **Save important conversations** for analysis and debugging

## Conclusion

AgentLib provides a powerful and flexible framework for building AI agents in R. Whether you need basic conversational AI, specialized medical writing assistance, or comprehensive data analytics capabilities, AgentLib has you covered.

Key takeaways:

- Start with `BaseAgent` for general-purpose AI assistance
- Use specialized agents (`MedicalWriterAgent`, `DataAnalyticsAgent`) for domain-specific tasks
- Extend functionality with custom tools
- Leverage batch processing for efficiency
- Monitor performance and manage resources carefully
- Save and share agent configurations for reproducibility

For more advanced examples and detailed documentation, see:

- `examples/basic_usage.R` - Comprehensive basic examples
- `examples/advanced_examples.R` - Advanced workflows and custom tools
- Individual function documentation via `?function_name`

Happy agent building!